GPU Acceleration of the Material Point Method

### 2019-02-11 pres/pres.tex
more math

\documentclass{beamer}
\usetheme{Malmoe}
\usecolortheme{seahorse}

@@

\subsection{The Pretty Strong but Mathematically Weak Formulation}
\begin{frame}
\frametitle{The Pretty Strong but Mathematically Weak Formulation}
\begin{align}
  \int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot  \Bigl[(_0\rho_0)(_0\boldsymbol{a}) - & {}_t\boldsymbol{f}^{\text{body}}\Bigr] d _0\boldsymbol{x}= \nonumber \\
										     &  \int _ { \partial \Omega ^ { t^n } } _t\boldsymbol{q} \cdot \boldsymbol{\sigma} d_t\boldsymbol{A} - \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}
\end{align}
\end{frame}
\section{A MPM Guide on GPGPU Pitfalls and Optimizations}

\subsection{c}

>>>>

\documentclass{beamer}
\usetheme{Malmoe}
\usecolortheme{seahorse}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{datavisualization}
\usetikzlibrary{datavisualization.formats.functions}

@@

\subsection{The Pretty Strong but Mathematically Weak Formulation}
\begin{frame}
\frametitle{The Pretty Strong but Mathematically Weak Formulation}

\textbf{Weak Formulation} (or Principle of Virtual Work):\\
Dot product equations with arbitrarily 'test functions' $\boldsymbol{q}$ and apply divergence theorem:
\begin{align}
\int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot  \Bigl[(_0\rho_0)(_0\boldsymbol{a}) - & {}_0\boldsymbol{f}^{\text{body}}{}_0^tJ\Bigr] d _0\boldsymbol{x}= \nonumber \\
										     &  \int _ { \partial \Omega ^ { t^n } } _t\boldsymbol{q} \cdot \boldsymbol{\sigma} d_t\boldsymbol{A} - \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}.
\end{align}
A strong solution is also a solution to the weak formulation. Leave out body forces(like gravity) and boundary condition (e.g. collisions) for now:
\begin{equation}
  \int _ { \Omega ^ { 0 } } _0\boldsymbol{q} \cdot (_0\rho_0)(_0\boldsymbol{a}) d _0\boldsymbol{x}= \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}  d_t\boldsymbol{x}.
\end{equation}
\end{frame}
\begin{frame}[fragile]
\textbf{Weak Derivative}:
\vfill
\begin{minipage}{0.49\textwidth}
  $y=|t|$ has weak derivative:
  \\\\
$ v = \left\{ \begin{array} { l l }
{ - 1 , } & { \text { if } t < 0 } \\
{ c , } & { \text { if } t = 0 } \\
{ 1 , } & { \text { if } t > 0 }
\end{array} \right.$
\\

\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{tikzpicture}
\datavisualization [school book axes,
		    visualize as line/.list={fx},
		    fx={style={color=blue}}
]

data [set=fx,format=function] {
      var x : interval [-1.5:1.5] samples 5;
      func y = abs(\value x);
      };
\end{tikzpicture}
\end{minipage}

\vfill
\begin{minipage}{0.49\textwidth}
  Heaviside step function has no weak derivate: \\\\
  $y= \left\{ \begin{array} { l l } { 0 , } & { \text { if } t < 0 } \\ { 1 , } & { \text { if } t \geq 0 } \end{array} \right.$
\end{minipage}
\begin{minipage}{0.4\textwidth}

\begin{tikzpicture}

  \datavisualization [school book axes,
		      visualize as line/.list={fx},
		      fx={style={color=blue}}]
data[set=fx] {
  x, y
  -1.5, 0
  0, 0
  0, 1
  1.5, 1
};
\end{tikzpicture}
\end{minipage}
\vfill
Allows for point loads, material discontinuities and more. \cite{bathe2006finite}
\end{frame}
\subsection{Discretization of Space and Time}
\begin{frame}
\frametitle{Discretization of Space and Time}
\textbf{Time discretization} with implicit midpoint scheme:
  \begin{equation}
    \frac { y ^ { n + 1 } - y ^ { n } } { \Delta t } = f ^ { n + \frac{1}{2} } = f \left( t ^ { n } + \frac{\Delta t}{2}, \frac{1}{2} y ^ { n } + \frac{1}{2} y ^ { n + 1 } \right)
  \end{equation}
\begin{itemize}
  \item \textbf{implicit} requires linear system solve $\Rightarrow$ more stable, larger time steps
  \item \textbf{midpoint} as it conserves gov. equations
\end{itemize}

  \begin{equation}
    \Rightarrow \int _ { \Omega ^ { t } } _t\boldsymbol{q} \cdot {}_t\rho(_t\boldsymbol{v}^{n+1}-{}_t\boldsymbol{v}^{n}) d _t\boldsymbol{x}= \int _ { \Omega ^ { t^n} } \nabla {}_t\boldsymbol{q} : \boldsymbol{\sigma}^{n+\frac{1}{2}}  d_t\boldsymbol{x}.
  \end{equation}
\end{frame}

\begin{frame}[fragile]
\textbf{Space Discretization} is done in a Galerkin/FEM fashion with grid based interpolants $w_i$ with limited support. Here dyadic products
\begin{equation}
  w_i(\boldsymbol{x}) =w(\boldsymbol{x}-\boldsymbol{x}_i) = w(\frac{1}{h}(x-x_i))w(\frac{1}{h}(y-y_i)w(\frac{1}{h}(z-z_i))
\end{equation}
of cubic b-splines suffice:
\begin{equation}
w ( x ) = \left\{ \begin{array} { l l } { \frac { 1 } { 2 } | x | ^ { 3 } - | x | ^ { 2 } + \frac { 2 } { 3 } } & { 0 \leqslant | x | < 1 } \\ { \frac { 1 } { 6 } ( 2 - | x | ) ^ { 3 } } & { 1 \leqslant | x | < 2 } \\ { 0 } & { 2 \leqslant | x | } \end{array}. \right
\end{equation}
\end{frame}
\begin{frame}

\begin{tikzpicture}[domain=0:1]
\begin{axis}
\addplot {x*x};
\end{axis}
\end{tikzpicture}

\end{frame}
\begin{frame}
Thus set in:
$$
{}_t\boldsymbol{q}(\boldsymbol{x},t^n) = \sum_i \boldsymbol{e}_iw_i(\boldsymbol{x}),
{}_t\boldsymbol{v}^{n(+1)}(\boldsymbol{x}) = \sum_j \boldsymbol{v}^{n(+1)}_jw_j(\boldsymbol{x}).
$$
Combine it with Numerical Integration where the particles function as quadrature points:
\end{frame}
\section{A MPM Guide on GPGPU Pitfalls and Optimizations}
\subsection{d}
